---
title: "Cross validation (Publication)"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warnings = FALSE,
                      message = FALSE,
                      comment = "#>",
                      #results = "hide",
                      digits = 4,
                      error = FALSE)

## clean the R environment
graphics.off()
rm(list = ls())
freshr::freshr()

## load packages
library(here, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(gtsummary, quietly = TRUE)
library(flextable, quietly = TRUE)

library(devtools, quietly = TRUE)
load_all()

## check the directory for the file
# here::dr_here()
here::set_here()

## the figure or results should be saved 
# paste0("foldername/Sfilename_workingresult_", 
#      Sys.Date(), ".filetype")
```

```{css, echo=FALSE}
.scroll-100 {
  max-height: 300px;
  max-width: 1000px;
  overflow-y: auto;
  background-color: inherit;
}
```

## Cross-validation for different $\kappa$ values larger brokenstick knots

Remember to change the kappa each time

The codes are in the file `16_mlm_kappa_cv.R` and `16_slm_kappa_cv.R`file

## $\kappa$ from 10 to 200 (currently 2024-01-24 19:22:05)

### The SLM

The setup for this PLM:

-   anchor \<- c(14, 30, 60, 90) including 14 as the anchor
-   bsk_knots \<- c(seq(10, 100, by = 10), seq(120, 300, by = 20)) extreme
-   kappa1 \<- seq(10, 200, by = 10), it may seems not big enough to see the pattern

```{r}
#| label: "kappa_load_15"

# load("figure/tsa_16_table_kappa1_cross_validation_2024-01-24 17:22:05.Rdata")
# load("figure/tsa_16_table_kappa2_cross_validation_2024-01-24 18:07:14.Rdata")
load("figure/tsa_15_table_kappa2_cross_validation_2024-01-24 18:02:57.Rdata")
load("figure/tsa_15_table_kappa1_cross_validation_2024-01-24 19:33:33.Rdata")

# map_dfr(result_kappa0, ~unlist(., recursive = FALSE))

kappa200_slm <- rbind(map_dfr(result_kappa1, ~unlist(., recursive = FALSE)),
                      map_dfr(result_kappa2, ~unlist(., recursive = FALSE))) %>%
  as.data.frame() %>%
  mutate(kappa = seq(10, 200, by = 10)) %>%
  dplyr::select(kappa, everything())

# map_dfr(result_kappa5, ~unlist(., recursive = FALSE))
```

The trade-off is not very obvious, the MSE is still around 50, but BIAS is hard to see. The coverage is not very good, but it is not very bad either.

not sure whether the result is due to the brokenstick model or not

```{r}
#| label: "kappa200_slm_plot_15"
#| fig-cap: Cross validation for kappa
#| fig-subcap-location: top
#| fig-width: 4
#| fig-height: 2
#| fig-subcap:
#| - "CV200_SLM_BIAS"
#| - "CV200_SLM_MSE"
#| - "CV200_SLM_CR50"
#| - "CV200_SLM_CR80"
#| - "CV200_SLM_CR90"
#| warning: false
#| layout-ncol: 2

kappa200_slm_bias <- kappa200_slm %>%
  ggplot() +
  geom_line(aes(x = kappa, y = bias), color = "darkgreen") 
kappa200_slm_mse <- kappa200_slm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = mse), color = "indianred") 
  # scale_y_continuous("Bias", 
  #   sec.axis = sec_axis(~ ., name = "MSE"))
kappa200_slm_cr50 <- kappa200_slm %>%
  ggplot() +
  geom_line(aes(x = kappa, y = coverage50), color = "blue1") 
kappa200_slm_cr80<- kappa200_slm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = coverage80), color = "blue3") 
kappa200_slm_cr90<- kappa200_slm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = coverage90), color = "blue4")


kappa200_slm_bias + theme_bw() + theme(legend.position = "none")
kappa200_slm_mse + theme_bw() + theme(legend.position = "none")
kappa200_slm_cr50 + theme_bw() + theme(legend.position = "none")
kappa200_slm_cr80 + theme_bw() + theme(legend.position = "none")
kappa200_slm_cr90 + theme_bw() + theme(legend.position = "none")
```

### The MLM

probably will use 50 to 60 as the final matches criterion.

```{r}
#| label: "kappa200_mlm_15"

load("figure/tsa_16_table_kappa1_cross_validation_2024-01-24 17:22:05.Rdata")
load("figure/tsa_16_table_kappa2_cross_validation_2024-01-24 18:07:14.Rdata")

kappa200_mlm <- rbind(map_dfr(result_kappa1, ~unlist(., recursive = FALSE)),
                      map_dfr(result_kappa2, ~unlist(., recursive = FALSE))) %>%
  as.data.frame() %>%
  mutate(kappa = seq(10, 200, by = 10)) %>%
  dplyr::select(kappa, everything())
```

```{r}
#| label: "kappa200_mlm plot_15"
#| fig-cap: Cross validation for kappa
#| fig-subcap-location: top
#| fig-width: 4
#| fig-height: 2
#| fig-subcap:
#| - "CV200_MLM_BIAS"
#| - "CV200_MLM_MSE"
#| - "CV200_MLM_CR50"
#| - "CV200_MLM_CR80"
#| - "CV200_MLM_CR90"
#| warning: false
#| layout-ncol: 2


kappa200_mlm_bias <- kappa200_mlm %>%
  ggplot() +
  geom_line(aes(x = kappa, y = bias), color = "darkgreen") 
kappa200_mlm_mse <- kappa200_mlm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = mse), color = "indianred") 
  # scale_y_continuous("Bias", 
  #   sec.axis = sec_axis(~ ., name = "MSE"))
kappa200_mlm_cr50 <- kappa200_mlm %>%
  ggplot() +
  geom_line(aes(x = kappa, y = coverage50), color = "blue1") 
kappa200_mlm_cr80<- kappa200_mlm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = coverage80), color = "blue3") 
kappa200_mlm_cr90<- kappa200_mlm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = coverage90), color = "blue4")

kappa200_mlm_bias + theme_bw() + theme(legend.position = "none")
kappa200_mlm_mse + theme_bw() + theme(legend.position = "none")
kappa200_mlm_cr50 + theme_bw() + theme(legend.position = "none")
kappa200_mlm_cr80 + theme_bw() + theme(legend.position = "none")
kappa200_mlm_cr90 + theme_bw() + theme(legend.position = "none")

```

## Cross-validation for different $\kappa$ values smaller brokenstick knots

### The SLM

The setup for this PLM:

-   anchor \<- c(14, 30, 60, 90) including 14 as the anchor
-   bsk_knots \<- c(50, 100, 150, 200) relatively reasonable
-   kappa1 \<- seq(10, 200, by = 10), it may seems not big enough to see the pattern

```{r}
#| label: "kappa_load_17"

# load("figure/tsa_16_table_kappa1_cross_validation_2024-01-24 17:22:05.Rdata")
# load("figure/tsa_16_table_kappa2_cross_validation_2024-01-24 18:07:14.Rdata")

load("figure/tsa_17_table_simple_bks_slm_cross_validation_2024-01-24 22:40:50.Rdata")

# map_dfr(result_kappa0, ~unlist(., recursive = FALSE))


kappa200_slm <- rbind(map_dfr(result_kappa1, ~unlist(., recursive = FALSE))) %>%
  as.data.frame() %>%
  mutate(kappa = seq(10, 200, by = 10)) %>%
  dplyr::select(kappa, everything())

kappa200_slm

# map_dfr(result_kappa5, ~unlist(., recursive = FALSE))
```

brokenstick model does affect the result, but not very much over the trend

```{r}
#| label: "kappa200_slm_plot_17"
#| fig-cap: Cross validation for kappa
#| fig-subcap-location: top
#| fig-width: 4
#| fig-height: 2
#| fig-subcap:
#| - "CV200_SLM_BIAS"
#| - "CV200_SLM_MSE"
#| - "CV200_SLM_CR50"
#| - "CV200_SLM_CR80"
#| - "CV200_SLM_CR90"
#| warning: false
#| layout-ncol: 2

kappa200_slm_bias <- kappa200_slm %>%
  ggplot() +
  geom_line(aes(x = kappa, y = bias), color = "darkgreen") 
kappa200_slm_mse <- kappa200_slm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = mse), color = "indianred") 
  # scale_y_continuous("Bias", 
  #   sec.axis = sec_axis(~ ., name = "MSE"))
kappa200_slm_cr50 <- kappa200_slm %>%
  ggplot() +
  geom_line(aes(x = kappa, y = coverage50), color = "blue1") 
kappa200_slm_cr80<- kappa200_slm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = coverage80), color = "blue3") 
kappa200_slm_cr90<- kappa200_slm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = coverage90), color = "blue4")


kappa200_slm_bias + theme_bw() + theme(legend.position = "none")
kappa200_slm_mse + theme_bw() + theme(legend.position = "none")
kappa200_slm_cr50 + theme_bw() + theme(legend.position = "none")
kappa200_slm_cr80 + theme_bw() + theme(legend.position = "none")
kappa200_slm_cr90 + theme_bw() + theme(legend.position = "none")

```

### The MLM

probably will use 50 to 60 as the final matches criterion.

```{r}
#| label: "kappa200_mlm_17"

load("figure/tsa_17_table_simple_bsk_mlm_cross_validation_2024-01-24 22:38:18.Rdata")

kappa200_mlm <- rbind(map_dfr(result_kappa2, ~unlist(., recursive = FALSE))) %>%
  as.data.frame() %>%
  mutate(kappa = seq(10, 200, by = 10)) %>%
  dplyr::select(kappa, everything())
```

```{r}
#| label: "kappa200_mlm plot_17"
#| fig-cap: Cross validation for kappa
#| fig-subcap-location: top
#| fig-width: 4
#| fig-height: 2
#| fig-subcap:
#| - "CV200_MLM_BIAS"
#| - "CV200_MLM_MSE"
#| - "CV200_MLM_CR50"
#| - "CV200_MLM_CR80"
#| - "CV200_MLM_CR90"
#| warning: false
#| layout-ncol: 2


kappa200_mlm_bias <- kappa200_mlm %>%
  ggplot() +
  geom_line(aes(x = kappa, y = bias), color = "darkgreen") 
kappa200_mlm_mse <- kappa200_mlm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = mse), color = "indianred") 
  # scale_y_continuous("Bias", 
  #   sec.axis = sec_axis(~ ., name = "MSE"))
kappa200_mlm_cr50 <- kappa200_mlm %>%
  ggplot() +
  geom_line(aes(x = kappa, y = coverage50), color = "blue1") 
kappa200_mlm_cr80<- kappa200_mlm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = coverage80), color = "blue3") 
kappa200_mlm_cr90<- kappa200_mlm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = coverage90), color = "blue4")
kappa200_mlm_bias + theme_bw() + theme(legend.position = "none")

kappa200_mlm_bias + theme_bw() + theme(legend.position = "none")
kappa200_mlm_mse + theme_bw() + theme(legend.position = "none")
kappa200_mlm_cr50 + theme_bw() + theme(legend.position = "none")
kappa200_mlm_cr80 + theme_bw() + theme(legend.position = "none")
kappa200_mlm_cr90 + theme_bw() + theme(legend.position = "none")

```

### SLM and MLM comparison

```{r}
#| label: "comparison_17"
#| fig-cap: Cross validation for kappa
#| fig-subcap-location: top
#| fig-width: 4
#| fig-height: 2
#| fig-subcap:
#| - "CV200_SLM_BIAS"
#| - "CV200_MLM_BIAS"
#| - "CV200_SLM_MSE"
#| - "CV200_MLM_MSE"
#| warning: false
#| layout-ncol: 2
kappa200_slm_bias + theme_bw() + theme(legend.position = "none") + ylim(0.45, 0.51)
kappa200_mlm_bias + theme_bw() + theme(legend.position = "none") + ylim(0.45, 0.51)

kappa200_slm_mse + theme_bw() + theme(legend.position = "none") + ylim(0.65, 0.85)
kappa200_mlm_mse + theme_bw() + theme(legend.position = "none") + ylim(0.65, 0.85)

# kappa200_mlm_cr50 + theme_bw() + theme(legend.position = "none")
# kappa200_mlm_cr80 + theme_bw() + theme(legend.position = "none")
# kappa200_mlm_cr90 + theme_bw() + theme(legend.position = "none")
```

Matching size of 50 seems a reasonable choice for the paper
