---
title: "Cross Validation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warnings = FALSE,
                      message = FALSE,
                      comment = "#>",
                      #results = "hide",
                      digits = 4,
                      error = FALSE)

## clean the R environment
# graphics.off()
# rm(list = ls())
# freshr::freshr()

## load packages
library(here, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(gtsummary, quietly = TRUE)
library(flextable, quietly = TRUE)

library(devtools, quietly = TRUE)
load_all()

## check the directory for the file
# here::dr_here()
here::set_here()

## the figure or results should be saved 
# paste0("foldername/Sfilename_workingresult_", 
#      Sys.Date(), ".filetype")
```

```{css, echo=FALSE}
.scroll-100 {
  max-height: 300px;
  max-width: 1000px;
  overflow-y: auto;
  background-color: inherit;
}
```

## Cross-validation for different $\kappa$ values larger brokenstick knots

Remember to change the kappa each time

The codes are in the file `16_mlm_kappa_cv.R` and `16_slm_kappa_cv.R`file

## $\kappa$ from 4 to 100 (currently 2024-01-24 19:22:05 update to 2024-04-04)

### The MLM {#sec-mlm_cv}

The setup for this PLM:

-   anchor \<- c(14, 30, 60, 90) including 14 as the anchor
-   bsk_knots \<- c(seq(10, 100, by = 10), seq(120, 300, by = 20)) extreme
-   kappa1 \<- seq(4, 100, by = 2), it may seems not big enough to see the pattern

```{r}
#| label: "kappa_load_15"

load("results/tsa_22_table_mlm_cv_2024-04-12 09:45:59.Rdata")

kappa200_mlm <- rbind(map_dfr(result_kappa1, ~unlist(., recursive = FALSE))) %>%
  as.data.frame() %>%
  mutate(kappa = seq(4, 200, by = 1)) %>%
  dplyr::select(kappa, everything())
# map_dfr(result_kappa5, ~unlist(., recursive = FALSE))
```

The trade-off is not very obvious, the MSE is still around 50, but BIAS is hard to see. The coverage is not very good, but it is not very bad either.

not sure whether the result is due to the brokenstick model or not

```{r}
#| label: "kappa200_mlm plot_15"
#| fig-cap: Cross validation for kappa
#| fig-subcap-location: top
#| fig-width: 4
#| fig-height: 2
#| fig-subcap:
#| - "CV_kappa200MLM_BIAS"
#| - "CV_kappa200MLM_MSE"
#| - "CV_kappa200MLM_CR50"
#| - "CV_kappa200MLM_CR80"
#| - "CV_kappa200MLM_CR90"
#| warning: false
#| layout-ncol: 2


kappa200_mlm_bias <- kappa200_mlm %>%
  ggplot() +
  geom_line(aes(x = kappa, y = bias), color = "darkgreen") 
kappa200_mlm_mse <- kappa200_mlm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = mse), color = "indianred") 
  # scale_y_continuous("Bias", 
  #   sec.axis = sec_axis(~ ., name = "MSE"))
kappa200_mlm_cr50 <- kappa200_mlm %>%
  ggplot() +
  geom_line(aes(x = kappa, y = coverage50), color = "blue1") 
kappa200_mlm_cr80<- kappa200_mlm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = coverage80), color = "blue3") 
kappa200_mlm_cr90<- kappa200_mlm %>%
  ggplot() +
  # geom_line(aes(x = kappa, y = bias), color = "blue") +
  geom_line(aes(x = kappa, y = coverage90), color = "blue4")

kappa200_mlm_bias + theme_bw() + theme(legend.position = "none")
kappa200_mlm_mse + theme_bw() + theme(legend.position = "none")
kappa200_mlm_cr50 + theme_bw() + theme(legend.position = "none")
kappa200_mlm_cr80 + theme_bw() + theme(legend.position = "none")
kappa200_mlm_cr90 + theme_bw() + theme(legend.position = "none")

```
